Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-07 paper/paper.tex
pass on section 3.1

\begin{figure}[h]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}

\Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
  Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
  respondam isto logo.}  \Gui{O window:64 seria removido
  sim.}\Mar{ok. vou explicar isto.}

  \Luc{O segundo campo da opção flags define flags que serão
    ignoradas. Nesse caso a regra procura por pacotes onde apenas a
    flag S está 'setada', independentemente do valor das flags 1 e
    2. Portanto a regra gerada pode capturar coisas diferentes da
    regra original. A inferencia do "12" n pode ser feita apenas a
    partir do pacote de entrada, precisariamos tratar
    isso.}\Mar{vc. pode explicar isto **mais detalhadamente** E (1)
    explicar que isto eh uma limitacao e (2) indicar a possivel consequencia
    disto?}

\tname{} only reports a rule if the \nids\ (\ie{}, Suricata)
instantiated with that rule is able to capture the traffic passed on
input. Consequently, by construction, a rule that uses this set of
options captures the malicious traffic. Unfortunately, this rule is
potentially overspecified. Informally, the rule includes ``too many''
characteristics of the malicious traffic used to define it and could
potentially result in false negatives.

Second, \tname{} uses negative (\ie{}, benign) traffic to address the
overfitting issue. \tname{} searches for alternative rules that
preserves the invariant that a rule should only capture the positive
traffic. It uses the rule obtained in the previous step to bootstrap
the search.

@@

@@

\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \caption{Suricata rule for Ping Scan.}
  \label{fig:synflood-example}
\end{figure}

>>>>

\begin{figure}[h]
  \vspace{-2ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-2ex}  
\end{figure}

%% \Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
%%   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
%%   respondam isto logo.}  \Gui{O window:64 seria removido
%%   sim.}\Mar{ok. vou explicar isto.}
  %% \Luc{O segundo campo da opção flags define flags que serão
  %%   ignoradas. Nesse caso a regra procura por pacotes onde apenas a
  %%   flag S está 'setada', independentemente do valor das flags 1 e
  %%   2. Portanto a regra gerada pode capturar coisas diferentes da
  %%   regra original. A inferencia do "12" n pode ser feita apenas a
  %%   partir do pacote de entrada, precisariamos tratar
  %%   isso.}\Mar{vc. pode explicar isto **mais detalhadamente** \Luc{a opção flags pode ter dois campos separados por virgula (ex: flags:S,12). o primeiro campo indica quais flags devem estar ativas em um pacote para que o suricata gere um alerta. se uma ou mais flags forem irrelevantes para a detecção de um determinado ataque (ou seja, se n importa se a flag está ou não ativa), elas serão adicionadas ao segundo campo.} E (1)
  %%   explicar que isto eh uma limitacao \Luc{n entendi exatamente oq explicar aqui.} e (2) indicar a possivel consequencia
  %%   disto? \Luc{a consequencia sao possiveis falsos negativos. ao não incluir flags no segundo campo podemos descartar pacotes por conta da presença de flags q n são relevantes (ou seja, vamos levar em conta flags q n deveriam interferir na detecção).}}

\noindent
\tname{} only reports a rule if the \nids\ (\eg{}, Suricata)
instantiated with that rule is able to capture the traffic passed on
input. Consequently, by construction, a rule that uses this set of
options captures the malicious traffic. Unfortunately, this rule is
potentially overspecified. Informally, the rule includes ``too many''
characteristics of the malicious traffic used to define it and could
potentially result in false negatives. In this particular case, note
that the option \CodeIn{windows: 64} is not included in the golden
rule.

Second, \tname{} uses negative (\ie{}, benign) traffic to address the
overfitting issue. \tname{} searches for alternative rules that
preserves the invariant that a rule should capture the positive
traffic but not capture the negative traffic. It uses the rule
obtained in the previous step to bootstrap the search and generates
various rules containing subsets of these three options.

@@

\noindent
\textbf{Differences.}~Note that there are differences between the
options produced in the first step of \tname{} and the options from
the golden rule, as per Figure~\ref{fig:synflood-example}. The option
\CodeIn{flags} does not mention parameters \CodeIn{1 2}, there is an
extra option \CodeIn{window}, and there are differences in the option
\CodeIn{threshold}. Considering the option \CodeIn{flags}, the first
parameter of that option indicates what flags in the packet must be
set while the second parameter indicates non-mandatory flags (\ie{},
``don't cares''). The option \CodeIn{flags: S} indicates that the flag
``S'' must be set and all flags not mentioned (``1'' and ``2'', in our
case) must be unset for the \nids\ to capture the
traffic. Consequently, that choice of option alone could result in
false negatives and false positives. An attack that satisfies all
other conditions in a rule and have flags ``1'' or ``2'' set would be
captured by the original rule but not the inferred rule---a case of
false negative. Likewise, an attack that satisfies all conditions and
have flags ``1'' and ``2'' unset would be captured by the inferred
rule but not the original rule---a case of false positive. The set of
positive examples is to blame in this case. To mitigate this issue,
the input messages should have covered cases with options ``1'' or
``2'' set. Considering option \CodeIn{windows: 64}, \tname{} is able
to generate rules without that option, as discussed above. Considering
option \CodeIn{threshold}, the values used in the golden rule are
domain-specific. \tname{} uses default values for these options. We
conjecture that the system administrator would be able to configure
this option.

@@

\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \caption{Suricata rule for Ping Scan.}
  \label{fig:pingscan-example}
\end{figure}
